---
title: "Payment for ecosystem services as a measure for tropical dry forest conservation in Colombia" 
subtitle: "Final work presented in Economics and Enviroment course"
author:
  name: "Juan Sebastián Mendoza Páez"
  affiliation: "Forest engineering student"
format:
  revealjs:
    footer: 'Payment for ecosystem services as a measure for tropical dry forest conservation in Colombia'
    center: true
    slide-number: true
    theme: Auxiliar files/Tema.scss
    self-contained: true
    margin: 0.1
    width: 1100
css: Auxiliar files/styles.css
bibliography: Auxiliar files/references.bib
csl: Auxiliar files/apa.csl
---

## Tropical dry forest {.title}


:::font-size="10px"
It is an ecosystem caracterized by: 

::: {.incremental}

- Strong rain stationality.
- Among 4 to 6 months per year of drought.
- Restricted to lowlands from Mexico to Bolivia and Brazil in neotropic.
- Physiological and morphological adaptations of plants and animals.
:::

:::

::: {.fragment}

:::{.callout-warning appearance="simple"}

::: justify
From ~9 million of hectares of original tropical dry forest extension, today only remain around of 8%. The livestock grazing and agriculture expansion are the main drivers of deforestation.
:::

:::

:::

::: cita
[@bst_2014] 
:::

## Payment for ecosystem services (PES) {.title}

::: justify
- **The ecosystem services** are the ecological characteristics, functions, or processes that directly or indirectly contribute to human wellbeing: that is, the benefits that people obtain from functioning ecosystems [@costanza2017].

::: {.incremental}
- **PES** is an economic instrument which aim is ensuring the provision of ecosystem services through direct and conditional incomes to population of strategic ecosystems.
:::

:::

## Objective {.title}

::: justify
The aim of this work was to determine the Colombian towns with potential to implement a PES scheme. The town potential was established through the opportunity cost^[The opportunity cost refers to the resources that people leave to perceive due to do not choice the best possible alternative.] of the main farming activities in regions where the tropical dry forest is present.
:::

## Methodology {.title}

![](Auxiliar files/methodology.svg)

## Results {.title}

::: grid

<div>
```{r}
pacman::p_load(leaflet, leaflet.extras, tidyverse, sf, geojsonsf, gt,
               gtExtras, formattable)

regiones <- st_read('Trabajo final/Shapes.gpkg', layer = 'regiones', quiet = T)
datos <- geojson_sf('Resultados.geojson') %>% 
  mutate(c_oportunidad = case_when(
          c_oportunidad == "Bajo" ~ "Low",
          c_oportunidad == "Medio" ~ "Medium",
          c_oportunidad == "Alto" ~ "High",
          c_oportunidad == "Muy Alto" ~ "Very High") %>%
            factor(levels = c('Low', 'Medium', 'High', 'Very High')),
        actividad = case_when(
          actividad == "cultivos" ~ "Crops",
          actividad == "forestal" ~ "Forestry",
          actividad == "ganaderia" ~ "Livestock",
          actividad == "avicultura" ~ "Poultri farming"
        ))

valores <- datos$c_oportunidad %>% unique

color <- c("#ff0000", "#ff8300", "#ffbd00", "#e7f529") %>% rev
paleta <- colorFactor(palette = color, valores)

texto <- paste0(
  "<b>Region:</b> ", datos$departamento %>% str_to_title, "<br/>",
  "<b>City:</b> ", datos$municipio %>% str_to_title, "<br/>",
  "<b>Activity:</b> ", datos$actividad, "<br/>",
  "<b>Opportunity cost/ha:</b> ", formattable::currency(datos$utilidad, symbol = '$'), "<br/>",
  "<b>Classification:</b> ", datos$c_oportunidad
) %>% lapply(htmltools::HTML)

leaflet(datos, width = 520) %>%
  addProviderTiles(provider = providers$CartoDB.VoyagerNoLabels) %>%
  addPolygons(data = regiones, weight = 0.5, group = 'Regions', fillOpacity = 0.2, col = 'gray') %>%
  addPolygons(weight = 1.5, color = ~paleta(datos$c_oportunidad),
              label = texto, group = 'Opportunity cost',
              labelOptions = labelOptions(textsize = "13px")) %>%
  addLegend(pal = paleta, values = ~c_oportunidad,
            opacity = 0.5, title = 'Opportunity cost', position = "topright") %>%
  addLayersControl(overlayGroups = c('Opportunity cost', 'Regions'),
                   options = layersControlOptions(collapsed = F)) %>%
  addMiniMap(position = "bottomright", toggleDisplay = T, width = 110, height = 120)
```
</div>

::: {style="display: flex; flex-direction: column; justify-content: center; font-size: 20px"}

::: {style="margin-bottom: 2vh"}
**Key findings:** 

- Andean region presented the major extension of tropical dry forest (~49%) followed by Orinoco region (~30%).

::: {.incremental}
- Approximately 300 thousands ha of tropical dry forest (30.1%) were classified with low opportunity cost.
:::

:::

::: {.fragment}

```{r}
read.csv('resumen.csv') %>% 
  mutate(region = case_when(
    region == 'andina' ~ 'Andean',
    region == "pacifica" ~ 'Pacific',
    region == 'caribe' ~ 'Caribbean',
    region == "orinoquia" ~ 'Orinoco'), 
    c_oportunidad = case_when(
          c_oportunidad == "Bajo" ~ "Low",
          c_oportunidad == "Medio" ~ "Medium",
          c_oportunidad == "Alto" ~ "High",
          c_oportunidad == "Muy Alto" ~ "Very High") %>%
            factor(levels = c('Low', 'Medium', 'High', 'Very High')),
    actividad = case_when(
          actividad == "cultivos" ~ "Crops",
          actividad == "forestal" ~ "Forestry",
          actividad == "ganaderia" ~ "Livestock",
          actividad == "avicultura" ~ "Poultri farming"
    )) %>%
  filter(c_oportunidad == "Low") %>%
  summarise(prop = n()/84*100, area_total = sum(area_bst),
            range = list(range(utilidad)), .by = region) %>%
  unnest_wider(col = 4) %>%
  rename('minimo' = 4, 'maximo' = 5) %>%
  gt() %>%
  gt_plt_bar_pct(column = prop, fill = 'blue', background = 'lightblue', scaled = T) %>%
  cols_align("center", contains("scale")) %>%
  cols_width(2 ~ px(120)) %>%
  gt_theme_nytimes() %>%
  tab_header("Low opportunity cost per hectare by Colombian regions") %>%
  cols_label(region = "Region", prop = "Municipalities",
                 area_total = 'Total Area (ha)', minimo = 'Minimum',
                 maximo = 'Maximum') %>%
  tab_options(column_labels.text_transform = "capitalize") %>%
  tab_spanner(label = 'Opportunity cost per hectare', columns = 4:5) %>%
  tab_style(
    style = list(
      cell_text(font = google_font(name = "Playfair Display"),
                    weight = '800', align = "left", color = '#203B46')),
    locations = cells_title(groups = "title")) %>%
  tab_style(
    style = list(
      cell_text(font = google_font(name = "Roboto Condensed"),
                    align = "center", size='small')),
    locations = cells_column_spanners()) %>%
  fmt_number(columns = 3) %>%
  fmt_currency(columns = 4:5)
```

:::

:::

:::

## Conclusions {.title}

- North departments of Colombia are key for the conservation of tropical dry forest. 

:::{.incremental}
- Opportunities to improve by getting information to fine scale.
:::

## Thanks {.title}

## References {.title}
